---
title: "Data Description"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}

library(tidyverse)

# draw in all datas


# election_cand

# imputed turnout

# ballot_precinct

# election_anc

# election_comms


f <- function(filepath){
    print(filepath)
    data = read.csv(filepath, sep=",", header=TRUE)
    head(data)
}



```

#### Terms:

Ward - division of the district, 1-8

ANC - Advisory Neighborhood Commission. Subdivision of ward identified by letter, e.g. 5D. Around 5-10 commissioners serve on an ANC.

SMD - "Single-member district"; subdivision of ANC from which a single ANC commissioner is elected


#### Format for cleaned_data filenames:

<year(s)>\_\<content>\_\<aggregation level>.csv

e.g.

2018_ancElection_commissioners_contest.csv is data from 2018 comprising both ANC Elections and subsequently incumbent commissioners, aggregated at the contest level







### 2012-2018 Election, contest-level



```{r, echo=FALSE}

f("../cleaned_data/2012_2018_ancElection_contest.csv")

```
- contest_name - ANC identifier
- smd_anc_votes - how many votes were cast in the election for this ANC commissioner
- explicit_candidates - how many official candidates were registered
- ward_ballots - ignore
- over_votes - number of ballots not counted due to multiple votes for ANC commissioner
- under_votes - number of ballots with no vote for ANC comm.
- ward_anc_votes - ignore
- anc
- smd
- ward - these are just contest_name broken out
- year
- winner
- winner_votes
- write_in_votes
- smd_ballots - number of ballots cast that could have included votes for this ANC election, even if they didnt



### 2012-2018 Elections, candidate-level

```{r, echo=FALSE}

f("../cleaned_data/2012_2018_ancElection_candidate.csv")

```


### 2012-2018 Elections, ANC-level

```{r, echo=FALSE}

f("../cleaned_data/2018_ancElection_anc.csv")

```

- ward, year, anc -- as above
- num_candidates -- average number of registered candidates
- votes - average number of votes for winning candidates in ANC
- vote_norm - average of ratio between winner votes and ballots cast (not total ANC votes)
- engagement - average of ratio between total ANC votes and ballots cast






### 2012-2018 imputed turnout, ANC-level

```{r, echo=FALSE}

f("../cleaned_data/2012_2018_imputedTurnout_anc.csv")

```

### 2012-2018 ballot totals, precinct-level

```{r, echo=FALSE}

f("../cleaned_data/2012_2018_ballots_precinct.csv")

```


### 2018 Elections and Incumbent Commissioners, contest-level


```{r, echo=FALSE}

f("../cleaned_data/2018_ancElection_commissioners_contest.csv")

```

produced by merge_incumbents.R

The first part of this dataset is just from 2012_2018_ancElection_contest.csv, but years before 2018 are filtered out.

- commissioner_name -- name of commissioner who occupied the seat following the 2018 election
- match - do the name of election winner and commissioner match? (subsequence-based similarity over .5)
- vacant - seat is vacant following election
- switcharoo - seat is recorded occupied by a person other than election winner
- substantive_write_in - write-in winner becomes named commissioner
- absent - named election winner with seat subsequently recorded vacant
- empty - write-in winner; seat vacant

